---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# flopsearch

<!-- badges: start -->
<!-- badges: end -->

flopsearch provides an effective and easy-to-use causal discovery algorithm for linear additive noise models.

## Installation

You can install the development version of flopsearch from [GitHub](https://github.com/) with:

``` r
install.packages("https://github.com/CausalDisco/flopsearch/releases/download/v0.1.3/flopsearch.tar.gz")
```

This requires a working installation of the [Rust toolchain](https://rust-lang.org/tools/install/).

The name of the installed package is ```flopsearch``` and it can be loaded with:

``` r
library(flopsearch)
```

## Citing FLOP
If you use FLOP in your scientific work, please cite this paper:
```bibtex
@article{cifly2025,
  author  = {Marcel Wien{"{o}}bst and Leonard Henckel and Sebastian Weichwald},
  title   = {{Embracing Discrete Search: A Reasonable Approach to Causal Structure Learning}},
  journal = {{arXiv preprint arXiv:2510.04970}},
  year    = {2025}
}
```

## Example

A simple example run of the FLOP algorithm provided by flopsearch.

```{r example}
library(flopsearch)

# set up data
p <- 10
W <- matrix(0, nrow = p, ncol = p)
W[cbind(1:(p-1), 2:p)] <- 1
X <- matrix(rnorm(10000 * p), nrow = 10000, ncol = p) %*% solve(diag(p) - W)
X_std <- scale(X)

# run flop
G <- flop(X, 2.0, restarts=50)
```

## Input and Output
As input, FLOP takes the data matrix, the BIC penalty parameter (we recommend ```2.0``` as a default choice) and either a ```timeout``` (in seconds) or the number of ILS restarts to control how long the search runs.

The output of FLOP is a CPDAG encoded with an adjacency matrix whose entry in row i and column j is 1 in case of a directed edge from the i-th to the j-th variable and 2 in case of an undirected edge between those variables (in case of an undirected edge, the entry in row j and column i is also 2, that is each undirected edge induce two 2's in the matrix).
